Estimation of flow trajectories in a multiple lines network

### 2022-04-25 presentation_bb/presentation_bb.tex
Presentation updated

\usepackage [autostyle, english = american]{csquotes}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}

@@

		Exact trajectories are impossible to know, but with additional hypotheses, we can \imp{estimate} them. \\
		\vspace{0.4cm}
		We will divide this problematic into two parts:
		\begin{itemize}
			\item The estimation of trajectories on a \imp{single line}.
			\item The estimation of trajectories on the \imp{multiple lines network}.
		\end{itemize}

@@

	\begin{frame}{Formal problem definition}
		Let a line (in one direction), which have $n$ stops. Let $\bm{\rho}_\text{in} = (\rho^\text{in}_s)$ and $\bm{\rho}_\text{out} = (\rho^\text{out}_t)$ be two vectors representing, respectively, the \imp{passengers entering and leaving lines at each stop}. \\
		\vspace{0.4cm}
		We search a $(n \times n)$ \imp{origin-destination matrix} $\mathbf{N} = (n_{st})$ where components represents
		$$
			n_{st} = \text{"\imp{the number of passengers entering line at $s$ and leaving at $t$}"}.
		$$
		These components must verify
		\begin{enumerate}
			\item $n_{st} \geq 0$,
			\item $n_{s\bullet} = \rho^\text{in}_s$,
			\item $n_{\bullet t} = \rho^\text{out}_t$.
		\end{enumerate}
		\small ($\bullet$ indicates a sum on the replaced index)
	\end{frame}

@@

	\begin{frame}{Formal problem definition}
		$$
		 \mathbf{N} = \bordermatrix{
			   & \sigma^\text{out}_1 & \sigma^\text{out}_2  & \cdots & \sigma^\text{out}_{n-1} & \sigma^\text{out}_n   \cr
			\sigma^\text{in}_1 & n_{11} & n_{12} & \cdots & n_{1,n-1} & n_{1n} \cr
			\sigma^\text{in}_2 & n_{21} & n_{22} & \cdots & n_{2,n-1} & n_{2n} \cr
			\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \cr
			\sigma^\text{in}_{n-1} & n_{n-1,1} & n_{n-1,2} & \cdots & n_{n-1,n-1} & n_{n-1,n} \cr
			\sigma^\text{in}_{n} & n_{n,1} & n_{n,2} & \cdots & n_{n,n-1} & n_{n,n}}
		$$
	\end{frame}

>>>>

\usepackage [autostyle, english = american]{csquotes}
\DeclareFontShape{OT1}{cmss}{b}{n}{<->ssub * cmss/bx/n}{} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}

@@

		Exact trajectories are impossible to know, but we can \imp{estimate} them. \\
		\vspace{0.4cm}
		We will divide this problematic into two parts:
		\begin{itemize}
			\item The estimation of trajectories on a \imp{single line}.
			\item The estimation of trajectories on the \imp{multiple lines network}.
		\end{itemize}

@@

	\begin{frame}{Formal problem definition}
		Let a line (in one direction), which have $n$ stops, indexed by line order. Let $\bm{\rho}_\text{in} = (\rho^\text{in}_s)$ and $\bm{\rho}_\text{out} = (\rho^\text{out}_t)$ be two vectors representing, respectively, the \imp{passengers entering and leaving lines at each stop}. \\
		\vspace{0.4cm}
		We search a $(n \times n)$ \imp{origin-destination matrix} $\mathbf{N} = (n_{st})$ where components represents
		$$
			n_{st} = \text{"\imp{the number of passengers entering line at $s$ and leaving at $t$}"}.
		$$
		These components must verify
		\begin{enumerate}
			\item $n_{st} \geq 0$,
			\item $n_{s\bullet} = \rho^\text{in}_s$,
			\item $n_{\bullet t} = \rho^\text{out}_t$.
		\end{enumerate}
		\small ($\bullet$ indicates a sum on the replaced index)
	\end{frame}

@@

	\begin{frame}{Formal problem definition}
		\small
		It reads:
		$$
		\mathbf{N} = \bordermatrix{
			& \rho^\text{out}_1 & \rho^\text{out}_2  & \cdots & \rho^\text{out}_{n-1} & \rho^\text{out}_n   \cr
			\rho^\text{in}_1 & n_{11} & n_{12} & \cdots & n_{1,n-1} & n_{1n} \cr
			\rho^\text{in}_2 & n_{21} & n_{22} & \cdots & n_{2,n-1} & n_{2n} \cr
			\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \cr
			\rho^\text{in}_{n-1} & n_{n-1,1} & n_{n-1,2} & \cdots & n_{n-1,n-1} & n_{n-1,n} \cr
			\rho^\text{in}_{n} & n_{n,1} & n_{n,2} & \cdots & n_{n,n-1} & n_{n,n}}
		$$

		In fact, we already know that some components are null:
		$$
		\mathbf{N} = \bordermatrix{
			& \imp{0} & \rho^\text{out}_2  & \cdots & \rho^\text{out}_{n-1} & \rho^\text{out}_n \cr
			\rho^\text{in}_1 & \imp{0} & n_{12} & \cdots & n_{1,n-1} & n_{1n} \cr
			\rho^\text{in}_2 & \imp{0} & \imp{0} & n_{23} & \cdots & n_{2n} \cr
			\vdots & \vdots & \vdots & \ddots & \ddots & \vdots \cr
			\rho^\text{in}_{n-1} & \imp{0} & \imp{0} & \cdots & \imp{0} & n_{n-1,n} \cr
			\imp{0} & \imp{0} & \imp{0} & \cdots & \imp{0} & \imp{0}}
		$$
	\end{frame}

	\begin{frame}{Formal problem definition}

		In this form, the problem is \imp{ill posed}, because it has multiple solutions. An exemple of solution is to make passengers follow a \imp{first in, first out (FIFO)} scheme. \\
		\vspace{0.4cm}
		A principle of mathematical modeling is to find the solution which makes the \imp{least assumptions about passenger behavior}, in other words the \imp{maximum entropy solution}. \\
		\vspace{0.4cm}
		In this case, it translates by supposing that there is the \imp{same probability of leaving the line for every passenger which have traveled at least one stop}.

	\begin{frame}{Solution with Markov chain modeling}

		We can then model passenger flow with a \imp{Markov chain}:

		\begin{center}
			\includegraphics[width=0.8\linewidth]{img/line_markov.png} 
		\end{center}

		with $p^\text{in}_i = \frac{\rho^\text{in}_i}{\rho^\text{in}_\bullet}$ and $p^\text{out}_i = \frac{\rho^\text{out}_i}{\sum_{1 \leq k \leq (i-1)} (\rho^\text{in}_k - \rho^\text{out}_k)}$.
	\end{frame}

	\begin{frame}{Solution with Markov chain modeling}

		\begin{center}
			\includegraphics[width=\textwidth]{img/stop_distributions.png} 
		\end{center}

	\begin{frame}{Solution with Markov chain modeling}

		\begin{center}
			\includegraphics[width=\textwidth]{img/length_exp.png} 
		\end{center}

	\section{Iterative proportial fitting}

	\begin{frame}{Iterative proportional fitting}

	The same solution can be obtained with the \imp{iterative proportional fitting (IPF)} algorithm. Let 
	\begin{enumerate}
		\item $\mathbf{P} = (p_{ij})$ a $(n \times m)$ matrix, 
		\item $\mathbf{u} = (u_i)$ a n-length vector, and 
		\item $\mathbf{v} = (v_i)$ a m-length vector,
	\end{enumerate}
	all of them with strictly positive components. We can find two vectors $\mathbf{a} = (a_i)$ and $\mathbf{b} = (b_i)$ such that the matrix $\mathbf{Q} = (q_{ij})$, defined with
	$$
		q_{ij} = a_i b_j p_{ij},
	$$
	verifies
	\begin{itemize}
		\item $q_{i\bullet} = u_i$,
		\item $q_{\bullet j} = v_j$,
		\item $K(\mathbf{Q}| \mathbf{P}) := \sum_{ij} \frac{q_{ij}}{q_{\bullet \bullet}} \log \left( \frac{q_{ij} / q_{\bullet \bullet}}{p_{ij} / p_{\bullet \bullet}} \right)$ is minimum.
	\end{itemize}
	\end{frame}

	\begin{frame}{Solution with iterative proportional fitting}

		In our context, it means that if we define an \imp{origin-destination affinity matrix} $\mathbf{S} = (s_{st})$ with 
			$$
				\mathbf{S} = \left( \begin{array}{ccccc}
				0 & 1 & \cdots & 1 & 1 \\
				0 & 0 & 1 & \cdots & 1 \\
				\vdots & \vdots & \ddots & \ddots & \vdots \\
				0 & 0 & \cdots & 0 & 1 \\
				0 & 0 & \cdots & 0 & 0 
				\end{array} \right)
			$$
		we can find the \imp{maximum entropy solution with iterative proportional fitting}, i.e we find $\mathbf{a} = (a_s)$ and $\mathbf{b} = (b_t)$, such that $n_{ij} = a_s b_t s_{st}$ verify:
		\begin{enumerate}
			\item $n_{s\bullet} = \rho^\text{in}_s$,
			\item $n_{\bullet t} = \rho^\text{out}_t$,
			\item $K(\mathbf{N}| \mathbf{S})$ is minimum.
		\end{enumerate}
		\small (a small number $\epsilon$ has to be added on null components).
	\end{frame}

	\begin{frame}{Solution with iterative proportional fitting}
		By \imp{decreasing (resp. increasing) $s_{st}$}, we \imp{reduce (resp. expand)} the resulting number of passengers going from $s$ to $t$ obtained with IPF. \\
		\vspace{0.4cm}
		Thus, this approach is more \imp{flexible}, because we could give a \imp{specific affinity matrix $\mathbf{S} = (s_{st})$}, based on other data (additional assumptions).\\
		\vspace{0.4cm}
		The relationship $n_{ij} = a_s b_t s_{st}$ can be seen as a \imp{gravity model} (still to investigate).
	\end{frame}

	\section{The multiple lines problem}
